\documentclass[]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{stmaryrd}

\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Primes}{\mathbb{P}}
\newcommand{\st}{\text{ s.t. }}
\newcommand{\txtand}{\text{ and }}
\newcommand{\txtor}{\text{ or }}
\newcommand{\lxor}{\veebar}

%opening
\title{Notes on Combinatorics}
\author{DSBA Mathematics Refresher 2024}
\date{}

\begin{document}
	
	\maketitle
	
	\begin{abstract}
		
	\end{abstract}
	
	
	Combinatorics is the branch of mathematics dealing with counting, arrangement, and combination of objects.
	It provides the foundation for various concepts in probability, statistics, computer science, and more.
	
	\paragraph{Choosing $n$ Times: Power $n$}
	\noindent\\
	When we choose an item from a set $n$ times \textit{\textbf{with replacement}}, each choice is independent of the previous ones.
	If the set has $k$ distinct elements, the total number of ways to make these choices is given by:
	$ k^n $
	
	\textbf{Example:}
	\noindent\\
	Suppose you have a set of 3 elements $\{a, b, c\}$.
	The number of different sequences of length 4 that can be formed by choosing from this set with replacement is:
	$ 3^4 = 81 $
	
	\paragraph{Ordering $n$ Different Items: Factorial}
	\noindent\\
	When selecting $n$ different items from a set, the number of ways to arrange these $n$ items in order is given by $n!$ (read as "n factorial"), where:
	$$
	n! = n \times (n-1) \times \cdots \times 2 \times 1
	$$
	
	\textbf{Example:}
	\noindent\\
	For a set of 5 distinct items $\{a, b, c, d, e\}$, the number of ways to arrange all 5 items is:
	$$
	5!
	= 5 \times 4 \times 3 \times 2 \times 1
	= 120
	$$
	
	\paragraph{Choosing $k$ from $n$ \textit{(in Order)}}
	\noindent\\
	When choosing $k$ elements from a set of $n$ elements (without replacement) where the order of selection matters (permutations), the number of possible arrangements is given by:
	$$
	P(n, k) = \frac{n!}{(n-k)!}
	$$
	
	\textbf{Example:}
	\noindent\\
	Consider a set of 6 elements $\{a, b, c, d, e, f\}$, and you want to select and arrange 3 elements. The number of different arrangements is:
	$$
	P(6, 3)
	= \frac{6!}{(6-3)!}
	= \frac{6 \times 5 \times 4 \times 3 \times 2 \times 1}{3 \times 2 \times 1}
	= 120
	$$
	
	\paragraph{Choosing $k$ from $n$ \textit{(Without Order)}}
	\noindent\\
	When choosing $k$ elements from a set of $n$ elements where the order of selection does not matter (combinations), the number of possible combinations is given by the binomial coefficient:
	$$
	\binom{n}{k} = \frac{n!}{k!(n-k)!}
	$$
	
	\textbf{Example:}
	\noindent\\
	Given a set of 7 elements $\{a, b, c, d, e, f, g\}$, the number of ways to choose 3 elements without regard to order is:
	$$
	\binom{7}{3}
	= \frac{7!}{3!(7-3)!}
	= \frac{7 \times 6 \times 5 \times 4 \times 3 \times 2 \times 1}{(3 \times 2 \times 1)(4 \times 3 \times 2 \times 1)}
	= 35
	$$
	
	\newpage
	\textbf{{\LARGE Extra: Fundamental Mathematics}}
	This part was improvised, as we went through the planned lecture (much) faster than expected.
	
	\section*{Modular Arithmetic}
	\noindent \\
	\textbf{Definition}: Modular arithmetic involves integers where numbers reset to zero after reaching a specified modulus.
	\noindent \\
	\textbf{Real-world Example}: The 12-hour clock system is a common example, where after 12, the next hour is 1.
	\noindent \\
	\textbf{Notation}: $a \equiv b \ (\text{mod} \ m)$
	This means that when $a$ is divided by $m$, it leaves a remainder $b$.

	
	\subparagraph{Basic Concepts}
	\noindent \\
	\textbf{Modulus (m)}: The value at which numbers wrap around.
	\noindent \\
	\textbf{Congruence Relation}: $a \equiv b \ (\text{mod} \ m)$ means $a - b$ is divisible by $m$.
	\noindent \\
	\textbf{Examples}:
	\begin{itemize}
		\item $17 \equiv 5 \ (\text{mod} \ 12)$ because $17 - 5 = 12$, which is divisible by 12.
		\item $20 \equiv 2 \ (\text{mod} \ 9)$ because $20 - 2 = 18$, which is divisible by 9.
	\end{itemize}

	\subparagraph{Exercises}
	Calculate the following:
	\begin{enumerate}
		\item $25 \ (\text{mod} \ 7)$
		\item $37 \ (\text{mod} \ 10)$
		\item $100 \ (\text{mod} \ 13)$
	\end{enumerate}
	
	\subparagraph{Basic Properties}
	\noindent \\
	\textbf{Addition}: If $a \equiv b \ (\text{mod} \ m)$ and $c \equiv d \ (\text{mod} \ m)$, then $(a + c) \equiv (b + d) \ (\text{mod} \ m)$.
	\noindent \\
	\textbf{Subtraction}: $(a - c) \equiv (b - d) \ (\text{mod} \ m)$.
	\noindent \\
	\textbf{Multiplication}: $(a \times c) \equiv (b \times d) \ (\text{mod} \ m)$.
	
	\subparagraph{Link with divisibility}
	\noindent \\
	Modular arithmetic is closely related to the concept of divisibility.
	Specifically, the congruence $a \equiv b \ (\text{mod} \ m)$ implies that the difference $a - b$ is divisible by $m$.
	In other words, there exists an integer $k$ such that $a - b = km$.
	This relationship shows that modular arithmetic can be used to test for divisibility.
	For example, a number $a$ is divisible by $m$ if and only if $a \equiv 0 \ (\text{mod} \ m)$.
	Thus, modular arithmetic provides a powerful tool for solving problems related to divisibility and for simplifying complex calculations by working with remainders.
	
	\subparagraph{Exercises}
	Simplify the following:
	\begin{enumerate}
		\item $(4 + 9) \ (\text{mod} \ 7)$
		\item $(5 \times 8) \ (\text{mod} \ 11)$
		\item Show that $2 \mid k^2 \implies 2 \mid k$ \textit{(hint: separate cases)}		\item Show that $3 \mid k^2 \implies 3 \mid k$ \textit{(hint: separate cases)}
	\end{enumerate}

	\vspace{3cm}
	\section*{Proof by Contradiction}
	\subsection*{$\sqrt{2}$ is not a rational (i.e.: $\sqrt{2} \not\in \mathbb{Q}$)}
	
	We will prove that $\sqrt{2}$ is not a rational number using a proof by contradiction.
	
	\begin{proof}
		Assume, for the sake of contradiction, that $\sqrt{2}$ is rational. Then, we can write $\sqrt{2}$ as a fraction of two integers $\frac{p}{q}$.
		In case $p$ and $q$ have common divisor $k > 1$, we can divide both by that $k$, untill $p$ and $q$ become coprime.
		Hence, we can now assume $\sqrt{2} = \frac{p}{q}$, where $p$ and $q$ are coprime (i.e., they have no common factors other than 1), and $q \neq 0$.
		Hence, we have:
		$$ \sqrt{2} = \frac{p}{q} $$
		Squaring both sides gives:
		$$ 2 = \frac{p^2}{q^2} $$
		Multiplying both sides by $q^2$ results in:
		$$ p^2 = 2q^2 $$
		This implies that $p^2$ is even, and therefore $p$ must also be even (property derived in the modular arithmetic section).
		So, we can write $p = 2k$ for some integer $k$.
		
		Substituting $p = 2k$ into the equation $p^2 = 2q^2$, we get:
		$$ (2k)^2 = 2q^2 $$
		so $$ 4k^2 = 2q^2 $$
		Dividing both sides by 2 yields:
		$$ 2k^2 = q^2 $$
		This implies that $q^2$ is even, and therefore $q$ must also be even.
		
		However, if both $p$ and $q$ are even, then $p$ and $q$ have a common factor of 2, which contradicts our initial assumption that $p$ and $q$ are coprime.
		
		Since our assumption that $\sqrt{2}$ is rational leads to a contradiction, we conclude that $\sqrt{2}$ is not rational.
		Therefore, $\sqrt{2} \not\in \mathbb{Q}$.
	\end{proof}
	
	\subparagraph{Exercises}
	Show, using the same technique that $\sqrt{3} \not\in \mathbb{Q}$.
	
\end{document}
